Monitoring forest health using hyperspectral imagery: Does feature selection improve the performance of machine-learning techniques?

### 2021-06-24 docs/00-manuscripts/ieee/ieee-manuscript.tex
optimize abstract

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
	This study analyzed highly-correlated, feature-rich datasets from hyperspectral remote sensing data using multiple machine and statistical-learning methods.
	The effect of filter-based feature-selection methods on predictive performance was compared.
	Also, the effect of multiple expert-based and data-driven feature sets, derived from the reflectance data, was investigated.
	Defoliation of trees (\%) was modeled as a function of reflectance, and variable importance was assessed using permutation-based feature importance.
	Overall support vector machine (SVM) outperformed others such as random forest (RF), extreme gradient boosting (XGBoost), lasso (L1) and ridge (L2) regression by at least three percentage points.
	The combination of certain feature sets showed small increases in predictive performance while no substantial differences between individual feature sets were observed.
	For some combinations of learners and feature sets, filter methods achieved better predictive performances than the unfiltered feature sets, while ensemble filters did not have a substantial impact on performance.
	Permutation-based feature importance estimated features around the red edge to be most important for the models.
	However, the presence of features in the near-infrared region (800 nm - 1000 nm) was essential to achieve the best performances.
	More training data and replication in similar benchmarking studies is needed for more generalizable conclusions.
	Filter methods have the potential to be helpful in high-dimensional situations and are able to improve the interpretation of feature effects in fitted models, which is an essential constraint in environmental modeling studies.

@@

Penalized methods showed promising performances, especially when taking runtime into account.
When removing features with a correlation of nearly 1, lasso and ridge are able to score performances around 31 p.p..

@@

The large differences of selected features per fold during tuning give interesting insights into internals of the used models (\autoref{tab:tune-perc-sel-features}).
Values of selected feature percentages like 99.97268 \% in \autoref{tab:tune-perc-sel-features} emerge from the specific setting which got elected as the best setting during the tuning step.
The value for the amount of features to select is initially defined by an internal surrogate learner of the MBO tuning method and can be any numeric value between 0 and 1.

While in most cases, SVM and XGBoost require a substantial portion of all available features to achieve robust predictions, RF was able to achieve the best results with a relatively small amount of features.
Realizing early that few features are needed during tuning to reach adequate performances can reduce the overall computational runtime substantially, especially when iterating over parameters such as $m_\textrm{try}$ whose optimum (and range) depends on the number of features.
Hence, regardless of the potential advantage of using filters for increased predictive performance, it should be noted that these can have a strong positive effect on runtime, at least for RF in this study.

@@

While most defoliation studies operate on the plot level using coarser-resolution multispectral satellite data \cite{townsend2012, debeurs2008, rengarajan2016}, there are also several recent studies using airborne or ground-based sensors at the tree level.
Among these, \cite{meng2018, kalin2019} used ground-level methods such as \ac{ALS} or \ac{LiDAR}.

Studies focusing on tree-level defoliation used ground-level methods such as \ac{ALS} or \ac{LiDAR} \cite{meng2018, kalin2019}.
\cite{meng2018} used \ac{OLS} regression methods while \cite{kalin2019} retrieved information from ground-level RGB photos using \ac{CNN}.
However, both did not use spatial \ac{CV} and \cite{kalin2019} no \ac{FS}.
\cite{goodbody2018} used a \ac{PLS} model with high-resolution \ac{DAP} to predict cumulative defoliation caused by the spruce budworm.
Study results indicated that spectral features were found to be most helpful for the model.
Incorporating such (both spectral and structural) could be a possible enhancement for future works.

% write about hyperspectral remote sensing
The field of (hyperspectral) remote sensing has a strong focus on using RF for modeling in recent years \cite{belgiu2016}.
However, in high-dimensional scenarios, tuning parameter \texttt{\(m_{try}\)} becomes computationally expensive.
To account for this and the high dimensionality in general, studies used feature selection approaches like semi-supervised feature extraction \cite{xia2015}, wrapper methods \cite{fassnacht2014a, feng2016, georganos2018}, PCA and adjusted feature selection \cite{rochac2016}.
However, no study that made use of filter methods in combination with hyperparameter tuning in the field of (hyperspectral) remote sensing could be found.
Potential reasons for this gap could be an easier access to wrapper methods and a higher general awareness of such compared to filter methods.
Applying the filter-based feature selection methodology shown in this study and its related code provided in the research compendium might be a helpful reference for future studies using hyperspectral remote sensing data.

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
	This study analyzed highly-correlated, feature-rich datasets from hyperspectral remote sensing data using multiple machine and statistical-learning methods.
	The effect of filter-based feature-selection methods on predictive performance was compared.
	Also, the effect of multiple expert-based and data-driven feature sets, derived from the reflectance data, was investigated.
	Defoliation of trees (\%), derived from in-situ measurements in fall 2016, was modeled as a function of reflectance.
	Variable importance was assessed using permutation-based feature importance.
	Overall support vector machine (SVM) outperformed other algorithms such as random forest (RF), extreme gradient boosting (XGBoost), lasso (L1) and ridge (L2) regression by at least three percentage points.
	The combination of certain feature sets showed small increases in predictive performance while no substantial differences between individual feature sets were observed.
	For some combinations of learners and feature sets, filter methods achieved better predictive performances than using no feature selection.
	Ensemble filters did not have a substantial impact on performance.
	Permutation-based feature importance estimated features around the red edge to be most important.
	However, the presence of features in the near-infrared region (800 nm - 1000 nm) was essential to achieve the overall best performances.
	More training data and replication in similar benchmarking studies is needed for more generalizable conclusions.
	Filter methods have the potential to be helpful in high-dimensional situations and are able to improve the interpretation of feature effects in fitted models, which is an essential constraint in environmental modeling studies.

@@

@@

Arbitrary seeming values of selected feature percentages like 99.97268 \% in \autoref{tab:tune-perc-sel-features} emerge from the specific setting which got elected as the best setting during the tuning step.
The specific value to benchmark during tuning is defined by the internal surrogate learner of the MBO tuning method in this study and can be any numeric value between 0 and 1 (interpreted as percentage of available features).

% filters always reduce computation time
Realizing early during hyperparameter optimization that only few features are needed to reach adequate performances can reduce the overall computational runtime substantially.
Hence, regardless of the potential advantage of using filters for increased predictive performance, these can have a strong positive effect on runtime, especially on models which make use of hyperparameters that depend on the available number of features, such as RF $m_\textrm{try}$.

@@

% defoliation studies and what data they used
While most defoliation studies operate on the plot level using coarser-resolution multispectral satellite data \cite{townsend2012, debeurs2008, rengarajan2016}, there are also several recent studies using airborne or ground-based sensors at the tree level.
Among these, \cite{meng2018, kalin2019} used ground-level methods such as \ac{ALS} or \ac{LiDAR}.

% defoliation studies and what
Studies focusing on tree-level defoliation mainly used ground-level methods such as \ac{ALS} or \ac{LiDAR} \cite{meng2018, kalin2019}.
\cite{meng2018} used \ac{OLS} regression methods while \cite{kalin2019} retrieved information from ground-level RGB photos using \ac{CNN}.
However, both did not use spatial \ac{CV} and \cite{kalin2019} no \ac{FS}.
\cite{goodbody2018} used a \ac{PLS} model with high-resolution \ac{DAP} to predict cumulative defoliation caused by the spruce budworm.
Study results indicated that spectral features were found to be most helpful for the model.
Incorporating such (both spectral and structural) could be a possible enhancement for future works.
No studies were found modeling defoliation caused by \textit{Diplodia sapinea (Fr.) Fuckel} with remote sensing data and most studies focused on describing the tree conditions based on local sampling \cite{hlebarska2018, kaya2019}.

% write about hyperspectral remote sensing
The field of (hyperspectral) remote sensing has a strong focus on using RF for modeling in recent years \cite{belgiu2016}.
However, in high-dimensional scenarios, tuning parameter \texttt{\(m_{try}\)} becomes computationally expensive.
To account for this and the high dimensionality in general, studies used feature selection approaches like semi-supervised feature extraction \cite{xia2015}, wrapper methods \cite{fassnacht2014a, feng2016, georganos2018}, PCA and adjusted feature selection \cite{rochac2016}.
In general, applying feature selection methods on hyperspectral datasets has shown to be effective, regardless of the method used \cite{pal2010, keller2016}.
However, no study that made explicit use of filter methods in combination with hyperparameter tuning in the field of (hyperspectral) remote sensing could be found.
Potential reasons for this absence could be an easier programmatic access to wrapper methods and a higher general awareness of such compared to filter methods.
Applying the filter-based feature selection methodology shown in this study and its related code provided in the research compendium might be a helpful reference for future studies using hyperspectral remote sensing data.
